---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity<- read.csv('activity.csv',sep = ",",header = TRUE, na.strings ="NA",
                   colClasses = c('integer','Date','factor'))
library(ggplot2)
library(dplyr)
str(activity)
activity<- na.omit(activity)

```


## What is mean total number of steps taken per day?
```{r}
total.steps <- tapply(activity$steps, activity$date, FUN=sum)
plot1<- ggplot(activity,aes(date, steps))+ geom_histogram(stat="identity", binwidth = 0.5)+ labs(title = "Histogram of Total Number of Steps Taken Each Day",x = "Date", y = "Total Number of Steps")
print(plot1)
```
Mean And Median of steps per day
```{r}
mean(total.steps)
median(total.steps)
```

## What is the average daily activity pattern?
```{r}
averages <- aggregate(activity$steps, list(interval = as.numeric(as.character(activity$interval))), FUN = "mean")
names(averages)[2] <- "Avg.Steps"

plot2 <- ggplot(averages, aes(interval, Avg.Steps)) + geom_line(color = "green", size = 0.7) + labs(title = "Time Series Plot of the 5-minute Intervals", x = "5-minute intervals", y = "Average Number of Steps Taken")
print(plot2)
```

Which 5-minute interval, on average across all days, contains the most steps?
```{r}
averages[averages$Avg.Steps == max(averages$Avg.Steps),]
```

Imputing missing values by using mean for that 5-minute interval.


## Imputing missing values
```{r}
impData <- activity 
for (i in 1:nrow(impData)) {
    if (is.na(impData$steps[i])) {
        impData$steps[i] <- averages[which(impData$interval[i] == averages$interval), ]$Avg.Steps
    }
}
sum(!complete.cases(impData))
plot3 <- ggplot(impData, aes(date, steps)) + geom_histogram(stat = "identity",binwidth = .5) +
        labs(title = "Histogram of Total Number of Steps Taken Each Day (Imputed Data)",x = "Date", y = "Total Number of Steps")
print(plot3)

total.steps.impute <- tapply(impData$steps, impData$date, FUN = sum)
mean(total.steps.impute)

median(total.steps.impute)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
impData$weekdays <- factor(format(impData$date, "%A"))
levels(impData$weekdays)

levels(impData$weekdays) <- list(weekday = c("Monday", "Tuesday",
                                             "Wednesday", 
                                             "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
levels(impData$weekdays)

table(impData$weekdays)

new.averages <- aggregate(impData$steps, 
                      list(interval = as.numeric(as.character(impData$interval)), 
                           weekdays = impData$weekdays),
                      FUN = "mean")
names(new.averages)[3] <- "meanOfSteps"
library(lattice)
plot4 <- xyplot(new.averages$meanOfSteps ~ new.averages$interval | new.averages$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
print(plot4)
```
